### 2022-09-29 Thesis/text/methoden.tex
final v2

The equipped sensors and easy simulation capabilities make the robot well-suited for using the Navigation2 stack and further development for behavior planning. The robot specifications are specified in table \ref{tab:turtlebot_spec}. The turtlebot will be used for the analysis of the current state of mobile navigation. Furthermore, the simulated turtlebot will be the platform where the behavior planning will be implemented and tested.

@@

\subsection{Navigation2}

The ROS Navigation2 Stack (Nav2) combines different packages that allow mobile robots to navigate from point A to point B. Navigation2 is the de-facto standard for mobile navigation with a wide range of supported robots. Supported robot types are holonomic, differential-drive, legged, and ackerman (car-like). For the mobile robot to make use of the Nav2 stack, it has to be set up in a certain way to be able to generate plans and execute commands in the right way. The Nav2 setup requires the mobile robot to possess a laser scan or point cloud sensor, an odometry source (such as an Inertial Measurement Unit (IMU) or wheel encoders), a map with information about free and occupied spaces (Occupancy Grid Map (OGM)),and a set of transformations for planning and navigation (see \ref{fig:nav_architecture}).

@@

The stack contains tools that save and load occupancy maps, localize the robot, plan paths, execute the path, provide cost maps, build behaviors, and execute recoveries \cite{macenski2020}. These packages are often integrating a Simultaneous Localization and Mapping (SLAM) method to build or enlarge maps. Path planning and execution capabilities, which correspond to the global and local planner described in section \ref{sec:Autonomous Driving Navigation Architectures}, are further supported by ready-to-use planning plugins that use A* and Dijkstra algorithms for global path planning and a dynamic window approach for path execution (local planning). The system sequencing is done in a hierarchical structure as a central behavior tree calls asynchronous actions from the respective planners after another, as seen in figure \ref{fig:nav_architecture}.

@@

The behavior tree depicted in figure \ref{fig:bt_nav} is the default Nav2 behavior and has a set of recovery behaviors already implemented (spin, wait, back up, clear map). These behaviors are executed if the robot can not make progress towards the set goal. Nav2 uses a node lifecycle management system to start, activate, deactivate, and shut down nodes in a controlled way. The lifecycle approach allows the system to monitor the system nodes' startups, executions, and failures.

@@

Examples of such events currently not handled appropriately are slipping wheels, orientation changes by external influences, or undetected, more specifically, undetectable obstacles. This can lead to erratic movement of the robot, as reactive behaviors and more deliberative behaviors compete for authority over the robot's movement. By design, to allow more intelligent robot behaviors, the deliberative behavior can override commands from reactive behaviors. However, the deliberative behavior's planning is based upon a false representation of reality, leading to incorrect commands.

The unsafe planning triggers reactive-type behaviors again to counteract the commands from the deliberative behavior. In a more severe scenario the reactive behaviors may not get activated at all, and the robot continues to drive despite the incorrect interpretation of the surroundings. This highly unsafe behavior requires a human operator to step in and restore the robot manually to full functionality. 
%\textit{Add a closing statement about that suitability of the Nav stack for achieving autonomy}

@@

The software requirements that improve the robot's behavior are derived from the current limitations of the standard ROS and Navigation2 setup for mobile robots described in the previous section. The functional requirements and their acceptance criteria are listed in table \ref{tab:fn_req}. The non-functional requirements are listed seperately in table \ref{tab:nonfn_req}. To better contextualize the context of the systems requirements, figure \ref{fig:use_case} depicts a Use-Case diagram with the needed functions to improve the system's autonomy.

@@

\begin{table}[ht]
	\centering
	\caption{Functional Requirements}
	\label{tab:fn_req}
	\renewcommand{\arraystretch}{1.5}
	\resizebox{0.9\textwidth}{!}{%
		\begin{tabular}{| m{0.08\textwidth} | m{0.13\textwidth}| m{0.09\textwidth} | m{0.60\textwidth}|} 
	 	\hline
	 	\textbf{Nr.} & \textbf{Name} & \textbf{Priority} & \textbf{Description/Acceptance Criteria} \\ 
	 	\hline
	 	fn$\_$req1 & Sensor Failure & High & The system detects sensor failure. Ensure that the system can restart sensors and decrease the speed during the time the sensor delivers limited information. \\ 
	 	\hline
	 	fn$\_$req2 & Emergency Detection & High & The system detects emergency. Ensure that the system can detect when the continuation on the calculated path is no longer safe (sensor failures, blockage). \\
	 	\hline
	 	fn$\_$req3 & Emergency Stop & High & The system can initiate emergency stops. Ensure that the system can override all commands and stop in case an emergency is detected. \\
	 	\hline 	
	 	fn$\_$req4 & Override Navigation2 & High & The system can override navigation2. Ensure that the system's commands can always override the commands coming from navigation2. \\
	 	\hline
	 	fn$\_$req5 & Maintain operability & High & The robot executes commands as long as it is safe. Ensure that the robot keeps driving if it is safe even when system functions are not working correctly. \\
	 	\hline
	 	fn$\_$req6 & Recovery & High & The system can recover from crashes. Ensure that the system can successfully reach goals despite previous crashes.\\ 
	 	\hline 
	 	fn$\_$req7 & Control Path Planning & Medium & The system controls and rates the quality of the planned paths. \\
	 	\hline	
	 	fn$\_$req8 & Reset Goals & Medium & The system can reset and override goals set by the user so that the goal is reachable by planners. \\
	 	\hline
	 	fn$\_$req9 & Robot Range & Medium & Ensure that the robot will not run out of battery during navigation to a goal. \\	
	 	\hline
		\end{tabular}
	}
\end{table}

@@

Additionally, a component to switch intelligently between speed commands from Nav2 and the autonomy layer is needed. This component could be realized as a decision gate that receives speed commands from Nav2 and the new behavior planning component and can forward, block or modify commands sent to the controller. This modification aspect allows high-quality local planning capabilities, but enables more deliberative behaviors to happen upon this planning. A simplified system block diagram is depicted in figure \ref{fig:block_diagram}. The proposed additional components, discussed in this chapter are colored in blue. Existing components, discussed in section \ref{sec:Current State}, are colored in gray.

@@

A comparison and assessment of the behavior planning approaches were made in section \ref{sec:Behavior Planning Approaches}. The decision which approach to use was made in favor of the behavior tree method. The main reason for this decision is the increased flexibility when creating and expanding the behavior planning compared to FSMs. Also, behavior trees allow full determinism and introspection during execution which is why the POMPD approach is not suitable as the main behavior planning approach for the specified requirements.

@@

Additionally to the types of control nodes that are mentioned in section \ref{subsec:Behavior Trees}, the library adds the concept of reactivity into the catalog of control nodes. Reactive sequences and reactive fallbacks differ from normal ones in handling nodes that return the state "Running". Instead of ticking the node again, the whole sequence restarts, which is very useful for continuously checking a condition and executing an asynchronous action node that gets halted when a condition is not met anymore (see table \ref{tab:control_nodes_bt}).

\begin{table}[ht]
	\centering
	\caption{Control Nodes in BT.CPP}
	\label{tab:control_nodes_bt}
	\renewcommand{\arraystretch}{1.5}
	\resizebox{0.9\textwidth}{!}{%
		\begin{tabular}{ | l | l | l |} 
%			{ | m{0.2\textwidth} | m{0.25\textwidth}| m{0.25\textwidth} | m{0.2\textwidth} |} 
	 	\hline
	 	\textbf{Name }& \textbf{Child returns Failure} & \textbf{Child Returns Running} \\ 
	 	\hline
	 	Sequence & Restart & Tick Again\\ 
	 	\hline
	 	Reactive Sequence & Restart & Restart \\ 
	 	\hline
	 	Sequence Star & Tick again & Tick again \\
	 	\hline
	 	Fallback & Tick next & Tick again\\
	 	\hline 	
	 	Reactive Fallback & Tick next & Restart \\
	 	\hline
		\end{tabular}
	}
\end{table}

Also, the library includes another class of Control Flow Nodes, called Decorator Nodes, which allow more control over the child node and its output. An important distinction to other Control Flow Nodes is that Decorator Nodes can only have one child node as compared to multiple children for sequence and fallback nodes. A comprehensive list of the available Decorators and descriptions is presented in table \ref{tab:decorators_bt}.

\begin{table}[ht]
	\centering
	\caption{Decorator Nodes in BT.CPP}
	\label{tab:decorators_bt}
	\renewcommand{\arraystretch}{1.5}
	\resizebox{0.9\textwidth}{!}{%
		\begin{tabular}{ | m{0.2\textwidth} | m{0.25\textwidth}| m{0.25\textwidth} | m{0.2\textwidth} |} 
	 	\hline
	 	\textbf{Name} & \textbf{Succeeds} & \textbf{Fails} & \textbf{Running} \\ 
	 	\hline
	 	Inverter Node & If the child returns false & If child return success & If the child returns running \\ 
	 	\hline
	 	ForceSuccess Node & Always & Never & If child returns running \\ 
	 	\hline
	 	ForceFailure Node & Never & Always & If child returns running\\
	 	\hline
	 	Repeat Node & Ticks child as long as it returns success. The maximum number of repeats can be defined & If the child returns failure & If the child returns running \\
	 	\hline 	
	 	Retry Node & If the child returns success & Ticks child as long as it returns failure. The maximum number of ticks can be defined & If the child returns running \\
	 	\hline
		\end{tabular}
	}
\end{table}

To allow the tree nodes to communicate, the library provides the developer with two possibilities. Either node can use the blackboard, a dictionary (key/value) that all nodes of the tree can read and write. Alternatively, two nodes can be connected through ports which allows direct communication between two nodes via a key/value.

@@

Groot is a software to create, edit, monitor, and debug behavior trees with a graphical user interface. The software allows the creation of behavior trees in XML files, which can be directly loaded into and used by the BT.CPP library. Groot can monitor the live execution of behavior trees and allows the introspection of how the tree reacts to different scenarios. Figure \ref{fig:groot} shows a small segment of an active behavior tree.

@@

In this graphical user interface, the condition nodes are labeled with a "C" before the node's name to distinguish them from action nodes which are tagged with an "A". The colors indicate the control flow and returned results of the respective nodes. A green box around a node shows that the tick signal got processed and returned "Success". A red box signals that the node returned "Failure". In the depicted example the "CollisionCheck" node returned "Failure" and the fallback node ticked a sequence node called "CollisionSequence". The control flow has not returned to the sequence yet, which is why the sequence's border is colored orange. The depicted tree is currently executing an action node at the bottom of the figure named "ReverseCmdVel".

@@

The development and testing of behaviors are done with the Gazebo simulator and the Turtlebot3 model. Gazebo is well-integrated into ROS and ROS2 and offers many ROS interfaces to control the simulation. Gazebo accurately models the physics of the robot and can simulate and publish the wheel odometry directly on a ROS topic. Furthermore, Gazebo has an integrated ODE physics engine and can simulate the readings for the laser scanner and IMU data via plugins. The use of the simulator allows easier repeatability for test cases as various obstacles can be spawned at any time, and sensor failures can be induced, resulting in quicker overall development.

Using a simulator facilitates to slow down or speed up the environment time. The time control allows more intricate observation of fast reactive type behaviors in slow motion while also enabling the observation of the long-time robustness of the system during extended tests. The ROS community offers many different simulation environments to test the robot. The environments range from simple geometric structures, to office spaces up to complete race tracks.

@@

The software versions used to implement and test the system on are listed below in table \ref{tab:versions}. Many of the software versions are determined by the choice of the ROS distribution (distro). ROS2 Foxy is not the newest ROS2 distribution but is well supported and developed, so the decision was made to use this distribution. The decision why the turtlebot is used is laid out in section \ref{subsec:Turtlebot3}. Although Python and C++ can be used to develop applications with ROS2, the behavior tree library relies on C++, which is why C++ was used for developing the whole system.

>>>>

The equipped sensors and easy simulation capabilities make the robot well-suited for using the Navigation2 stack and further development for behavior planning. The robot specifications are specified in table \ref{tab:turtlebot_spec} (p.\pageref{tab:turtlebot_spec}). The turtlebot will be used to analyze the current state of mobile navigation. Furthermore, the simulated turtlebot will be the platform where the behavior planning will be implemented and tested.

@@

\subsection{Navigation2}
\label{subsec:nav2}

The ROS Navigation2 Stack (Nav2) combines different packages that allow mobile robots to navigate from point A to point B. Navigation2 is the de-facto standard for mobile navigation with a wide range of supported robots. Supported robot types are holonomic, differential-drive, legged, and ackerman (car-like). For the mobile robot to make use of the Nav2 stack, it has to be set up in a certain way to be able to generate plans and execute commands in the right way. The Nav2 setup requires the mobile robot to possess a laser scan or point cloud sensor, an odometry source (such as an Inertial Measurement Unit (IMU) or wheel encoders), a map with information about free and occupied spaces (Occupancy Grid Map (OGM)), and a set of transformations for planning and navigation (see figure \ref{fig:nav_architecture}).

@@

The stack contains tools that save and load occupancy maps, localize the robot, plan paths, execute the path, provide cost maps, build behaviors, and execute recoveries \cite{macenski2020}. These packages are often integrating a Simultaneous Localization and Mapping (SLAM) method to build or enlarge maps. Path planning and execution capabilities, which correspond to the global and local planner described in section \ref{sec:Autonomous Driving Navigation Architectures}, are further supported by ready-to-use planning plugins that use A* and Dijkstra algorithms for global path planning and a dynamic window approach for path execution (local planning). The system sequencing is done in a hierarchical structure as a central behavior tree calls asynchronous actions from the respective planners after another, as seen in figure \ref{fig:nav_architecture} on page \pageref{fig:nav_architecture}.

@@

The behavior tree depicted in figure \ref{fig:bt_nav} is the default Nav2 behavior and has a set of recovery behaviors already implemented (spin, wait, back up, clear map). These behaviors are executed if the robot can not progress towards the set goal. Nav2 uses a node lifecycle management system to start, activate, deactivate, and shut down nodes in a controlled way. The lifecycle approach allows the system to monitor the system nodes' startups, executions, and failures.

@@

Examples of such events currently not handled appropriately are slipping wheels, orientation changes by external influences, or undetected, more specifically, undetectable obstacles. These events can lead to erratic movement of the robot, as reactive behaviors and more deliberative behaviors compete for authority over the robot's movement. By design, to allow more intelligent robot behaviors, the deliberative behavior can override commands from reactive behaviors. However, the deliberative behavior's planning is based upon a false representation of reality, leading to incorrect commands.

The unsafe planning triggers reactive-type behaviors again to counteract the commands from the deliberative behavior. In a more severe scenario, the reactive behaviors may not get activated, and the robot continues to drive despite the incorrect interpretation of the surroundings. This highly unsafe behavior requires a human operator to step in and restore the robot manually to full functionality. 
%\textit{Add a closing statement about that suitability of the Nav stack for achieving autonomy}

@@

The software requirements that improve the robot's behavior are derived from the current limitations of the standard ROS and Navigation2 setup for mobile robots described in the previous section. The non-functional requirements are listed separately in table \ref{tab:nonfn_req}. The functional requirements and their acceptance criteria are listed in table \ref{tab:fn_req} on page \pageref{tab:fn_req}. Figure \ref{fig:use_case} on page \pageref{fig:use_case} depicts a Use-Case diagram with the functions needed to improve the system's autonomy and better contextualize the system's requirements.

@@

\begin{table}[ht]
	\centering
	\caption{Functional Requirements}
	\label{tab:fn_req}
	\renewcommand{\arraystretch}{1.5}
	\resizebox{0.95\textwidth}{!}{%
		\begin{tabular}{| m{0.08\textwidth} | m{0.15\textwidth}| m{0.09\textwidth} | m{0.60\textwidth}|} 
			\hline
			\textbf{Nr.} & \textbf{Name} & \textbf{Priority} & \textbf{Description/Acceptance Criteria} \\ 
			\hline
			fn$\_$req1 & Sensor Failure & High & The system detects sensor failure. Ensure that the system can restart sensors and decrease the speed during the time the sensor delivers limited information. \\ 
			\hline
			fn$\_$req2 & Emergency Detection & High & The system detects emergency. Ensure that the system can detect when the continuation on the calculated path is no longer safe (sensor failures, blockage). \\
			\hline
			fn$\_$req3 & Emergency Stop & High & The system can initiate emergency stops. Ensure that the system can override all commands and stop in case an emergency is detected. \\
			\hline 	
			fn$\_$req4 & Override Navigation2 & High & The system can override navigation2. Ensure that the system's commands can always override the commands coming from navigation2. \\
			\hline
			fn$\_$req5 & Maintain operability & High & The robot executes commands as long as it is safe. Ensure that the robot keeps driving if it is safe even when system functions are not working correctly. \\
			\hline
			fn$\_$req6 & Recovery & High & The system can recover from crashes. Ensure that the system can successfully reach goals despite previous crashes.\\ 
			\hline 
			fn$\_$req7 & Control Path Planning & Medium & The system controls and rates the quality of the planned paths. \\
			\hline	
			fn$\_$req8 & Reset Goals & Medium & The system can reset and override goals set by the user so that the goal is reachable by planners. \\
			\hline
			fn$\_$req9 & Robot Range & Medium & Ensure that the robot will not run out of battery during navigation to a goal. \\	
			\hline
		\end{tabular}
	}
\end{table}

The prioritization is based on the severity of the consequences for the system. The functional requirements one to six focus on safety and robustness of the robot and are therefore prioritized high. Functional requirements seven to nine focus on extending the autonomy of the robot, but a failure in these requirements do not pose as high of a risk the other requirements.

@@

Additionally, a component to switch intelligently between speed commands from Nav2 and the autonomy layer is needed. This component could be realized as a decision gate that receives speed commands from Nav2 and the new behavior planning component and can forward, block or modify commands sent to the controller. This modification aspect allows high-quality local planning capabilities but enables more deliberative behaviors to happen upon this planning. A simplified system block diagram is depicted in figure \ref{fig:block_diagram}. The proposed additional components discussed in this chapter are colored in blue. Existing components, discussed in section \ref{sec:Current State}, are colored in gray.

@@

A comparison and assessment of the behavior planning approaches were made in section \ref{sec:Behavior Planning Approaches}. The decision on which approach to use was made in favor of the behavior tree method. The main reason for this decision is the increased flexibility when creating and expanding the behavior planning compared to FSMs. Also, behavior trees allow complete determinism and introspection during execution which is why the POMPD approach is unsuitable as the primary behavior planning approach for the specified requirements.

@@

Additionally to the types of control nodes mentioned in section \ref{subsec:Behavior Trees}, the library adds the concept of reactivity into the catalog of control nodes. Reactive sequences and reactive fallbacks differ from normal ones in handling nodes that return the state "Running". Instead of ticking the node again, the whole sequence restarts, which is very useful for continuously checking a condition and executing an asynchronous action node that gets halted when a condition is not met anymore (see table \ref{tab:control_nodes_bt}).

\begin{table}[ht]
	\centering
	\caption{Control Nodes in BT.CPP}
	\label{tab:control_nodes_bt}
	\renewcommand{\arraystretch}{1.5}
	\resizebox{0.8\textwidth}{!}{%
		\begin{tabular}{ | l | l | l |} 
			%			{ | m{0.2\textwidth} | m{0.25\textwidth}| m{0.25\textwidth} | m{0.2\textwidth} |} 
			\hline
			\textbf{Name }& \textbf{Child returns Failure} & \textbf{Child Returns Running} \\ 
			\hline
			Sequence & Restart & Tick Again\\ 
			\hline
			Reactive Sequence & Restart & Restart \\ 
			\hline
			Sequence Star & Tick again & Tick again \\
			\hline
			Fallback & Tick next & Tick again\\
			\hline 	
			Reactive Fallback & Tick next & Restart \\
			\hline
		\end{tabular}
	}
\end{table}

Also, the library includes another class of Control Flow Nodes, called Decorator Nodes, which allow more control over the child node and its output. An essential distinction to other Control Flow Nodes is that Decorator Nodes can only have one child node compared to multiple children for sequence and fallback nodes. A comprehensive list of the available Decorators and descriptions is presented in table \ref{tab:decorators_bt}.

\begin{table}[ht]
	\centering
	\caption{Decorator Nodes in BT.CPP}
	\label{tab:decorators_bt}
	\renewcommand{\arraystretch}{1.5}
	\resizebox{0.95\textwidth}{!}{%
		\begin{tabular}{ | m{0.15\textwidth} | m{0.4\textwidth}| m{0.4\textwidth} | m{0.2\textwidth} |} 
			\hline
			\textbf{Name} & \textbf{Succeeds} & \textbf{Fails} & \textbf{Running} \\ 
			\hline
			Inverter Node & If the child returns false & If child return success & If the child returns running \\ 
			\hline
			ForceSuccess Node & Always & Never & If child returns running \\ 
			\hline
			ForceFailure Node & Never & Always & If child returns running\\
			\hline
			Repeat Node & Ticks child as long as it returns success. Number of repeats can be defined & If the child returns failure & If the child returns running \\
			\hline 	
			Retry Node & If the child returns success & Ticks child as long as it returns failure. Number of ticks can be defined & If the child returns running \\
			\hline
		\end{tabular}
	}
\end{table}

To allow the tree nodes to communicate, the library provides the developer with two possibilities. Either node can use the blackboard, a dictionary (key/value) that all tree nodes can read and write. Alternatively, two nodes can be connected through ports which allows direct communication between two nodes via a key/value.

@@

Groot is a program to create, edit, monitor, and debug behavior trees with a graphical user interface. The software allows the creation of behavior trees in XML files, which can be directly loaded into and used by the BT.CPP library. Groot can monitor the live execution of behavior trees and allows the introspection of how the tree reacts to different scenarios. Figure \ref{fig:groot} shows a small segment of an active behavior tree.

@@

In this graphical user interface, the condition nodes are labeled with a "C" before the node's name to distinguish them from action nodes which are tagged with an "A". The colors indicate the control flow and returned results of the respective nodes. A green box around a node shows that the tick signal got processed and returned "Success", and a red box signals that the node returned "Failure". In the depicted example, the "CollisionCheck" node returned "Failure," and the fallback node ticked a sequence node called "CollisionSequence". The control flow has not returned to the sequence yet, which is why the sequence's border is colored orange. The depicted tree is executing an action node at the bottom of the figure named "ReverseCmdVel".

@@

The development and testing of behaviors are done with the Gazebo simulator and the Turtlebot3 model. Gazebo is well-integrated into ROS and ROS2 and offers many ROS interfaces to control the simulation. Gazebo accurately models the physics of the robot and can simulate and publish the wheel odometry directly on a ROS topic. Furthermore, Gazebo has an integrated ODE physics engine and can simulate the readings for the laser scanner and IMU data via plugins. The simulator allows easier repeatability for test cases as various obstacles can be spawned at any time, and sensor failures can be induced, resulting in quicker overall development.

Using a simulator facilitates slowing down or speeding up the environment time. The time control allows more intricate observation of fast reactive type behaviors in slow motion while also enabling the observation of the long-time robustness of the system during extended tests. The ROS community offers many different simulation environments to test the robot. The environments range from simple geometric structures to office spaces up to complete race tracks.

@@

The software versions used to implement and test the system are listed below in table \ref{tab:versions}. Many of the software versions are determined by the choice of the ROS distribution (distro). ROS2 Foxy is not the newest ROS2 distribution but is well supported and developed, so the decision was made to use this distribution. The decision why the turtlebot is used is laid out in section \ref{subsec:Turtlebot3}. Although Python and C++ can be used to develop applications with ROS2, the behavior tree library relies on C++, which is why C++ was used for developing the whole system.
